---
title: "Practica.TecnicasdeCampo.MasterBiologiaIntegrada.UniversidadSevilla"
author: "Marcial"
date: "05/21/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
1. Install and load packages

```{r}
#install.packages("picante")
#This will install also package dependences
library(picante)
#This will also load package dependences
#library(ape)
#library(vegan)
#library(permute)
#library(lattice)
#library(nlme)

#install.packages("brranching")
library(brranching)

#devtools::install_github("jinyizju/V.PhyloMaker")
library(V.PhyloMaker)

```


3. Frequency matrix.

```{r}
mydata <- read.csv("Transectos.csv", sep = ",")
mydata
mydata <- mydata[1:48,]
mydata

```



3. Create your tree and plot

```{r}

?phylomatic
###FIRST SOME EXAMPLE OF HOW TO GET A PHYLOGENY USING A PHYLOMATIC APPROACH

## Not run:
# Input taxonomic names
taxa <- c("Poa annua", "Phlox diffusa", "Helianthus annuus")
tree <- phylomatic(taxa=taxa, get = 'POST')
plot(tree, no.margin=TRUE)

# Genus names
taxa <- c("Poa", "Phlox", "Helianthus")
tree <- phylomatic(taxa=taxa, storedtree='R20120829', get='POST')
plot(tree, no.margin=TRUE)

# Lots of names
taxa <- c("Poa annua", "Collomia grandiflora", "Lilium lankongense", "Phlox diffusa", "Iteadaphne caudata", "Gagea sarmentosa", "Helianthus annuus")
tree <- phylomatic(taxa=taxa, get ='POST')
plot(tree, no.margin=TRUE)

# Don't clean - clean=TRUE is default(tree <- phylomatic(taxa=taxa, clean = FALSE))
## with clean=FALSE, you can get non-splitting nodes, which you
## need to collpase before plotting
library('ape')
plot(collapse.singles(tree), no.margin=TRUE)
# Output NeXML format
taxa <- c("Gonocarpus leptothecus", "Gonocarpus leptothecus", "Lilium lankongense")
out <- phylomatic(taxa=taxa, get ='POST', outformat = "nexml")
cat(out)
# Lots of names, note that when you have enough names (number depends on length of individual
# names, so there's no per se rule), you will get an error when using \code{get='GET'},
# when that happens use \code{get='POST'}
library("taxize")
spp <- names_list("species", 500)
# phylomatic(taxa = spp, get = "GET")
(out <- phylomatic(taxa = spp, get = "POST"))
plot(out)
# Pass in a tree from a URL on the web
spp <- c("Abies_nordmanniana", "Abies_bornmuelleriana", "Abies_cilicica", "Abies_cephalonica",
"Abies_numidica", "Abies_pinsapo", "Abies_alba")
url <- "http://datadryad.org/bitstream/handle/10255/dryad.8791/final_tree.tre?sequence=1"
phylomatic(taxa=spp, treeuri=url)


#Here we are going to create species tree of the species sampled during field work
### .....


taxa <- as.character(mydata[1:48,1])
mytree <- phylomatic(taxa=taxa, get = 'POST')
plot(mytree, no.margin=TRUE, cex = 0.8)

mytree2 <- phylomatic(taxa=taxa, get = 'POST', storedtree = "zanne2014")
plot(mytree2, no.margin=TRUE, cex = 0.8)


#################
#This is a new method recently published
taxa[6] <- "Clinopodium sp"
genera<-sapply(strsplit(taxa," "),function(x) x[1])
family <- c("Fabaceae", "Fabaceae", "Ericaceae", "Lamiaceae", "Asparagaceae", "Lamiaceae", "Fabaceae", "Ericaceae", "Cistaceae", "Cistaceae", "Cistaceae", "Rosaceae", "Fabaceae", "Thymelaeaceae", "Dioscoreaceae", "Boraginaceae", "Ericaceae", "Ericaceae", "Ericaceae", "Fabaceae", "Boraginaceae", "Asteraceae", "Lamiaceae", "Lamiaceae", "Myrtaceae", "Oleaceae", "Oleaceae", "Lamiaceae", "Asteraceae", "Anacardiaceae", "Rubiaceae", "Fagaceae", "Fagaceae", "Fagaceae", "Fagaceae", "Rhamnaceae", "Fabaceae", "Rhamnaceae", "Rosaceae", "Asparagaceae", "Scrophulariaceae", "Crassulaceae", "Smilacaceae", "Fabaceae", "Lamiaceae", "Lamiaceae", "Fabaceae", "Apocynaceae")
data <- data.frame(species = taxa, genus = genera, family = family)
my.phylo.maker <- phylo.maker(data, tree = GBOTB.extended, nodes = nodes.info.1, output.sp.list = TRUE, output.tree = FALSE, scenarios = c("S1","S2","S3"), r = 1)

plot(my.phylo.maker$scenario.3, cex = 0.6)

write.tree(my.phylo.maker$scenario.3, file = "mytree.tree")

mytree <- my.phylo.maker$scenario.3

```


4. Calculate phylogenetic diversity statistics.

```{r}
mydata[,1] <- chartr(" ", "_", mydata[,1])
mydata[6,1] <- "Clinopodium_sp"
setdiff(mydata[,1], mytree$tip.label)
setdiff(mytree$tip.label,mydata[,1])

class(mytree) <- "phylo"

color.plot.phylo(mytree, mydata, "Caliza_1", "Specie", leg.cex = 0.6, main = "Caliza_1") 
color.plot.phylo(mytree, mydata, "Caliza_2","Specie", leg.cex = 0.6, main = "Caliza_2") 
color.plot.phylo(mytree, mydata, "Dolomia_1", "Specie", leg.cex = 0.6, main = "Dolomia_1") 
color.plot.phylo(mytree, mydata, "Dolomia_2", "Specie", leg.cex = 0.6, main = "Dolomia_2") 
color.plot.phylo(mytree, mydata, "Brezal_1", "Specie", leg.cex = 0.6, main = "Brezal_1") 
color.plot.phylo(mytree, mydata, "Brezal_2", "Specie", leg.cex = 0.6, main = "Brezal_2") 
color.plot.phylo(mytree, mydata, "Brezal_3", "Specie", leg.cex = 0.6, main = "Brezal_3") 
color.plot.phylo(mytree, mydata, "Brezal_4", "Specie", leg.cex = 0.6, main = "Brezal_4") 
color.plot.phylo(mytree, mydata, "Alcornocal_1", "Specie", leg.cex = 0.6, main = "Alcornocal_1") 
color.plot.phylo(mytree, mydata, "Alcornocal_2", "Specie", leg.cex = 0.6, main = "Alcornocal_2") 
color.plot.phylo(mytree, mydata, "Alcornocal_3", "Specie", leg.cex = 0.6, main = "Alcornocal_3") 
color.plot.phylo(mytree, mydata, "Quejigal_1", "Specie", leg.cex = 0.6, main = "Quejigal_1") 
color.plot.phylo(mytree, mydata, "Quejigal_2", "Specie", leg.cex = 0.6, main = "Quejigal_2") 
color.plot.phylo(mytree, mydata, "Quejigal_3", "Specie", leg.cex = 0.6, main = "Quejigal_3") 


?pd
#Explore the function and run the examples

rownames(mydata) = mydata[,1]

mypd <- pd(t(mydata[,-c(1)]), mytree, include.root=TRUE)
mypd

?mpd
#Explore the function and run the examples

mympd <- mpd(t(mydata[,-c(1)]), cophenetic(mytree), abundance.weighted=T)
mympd
myresults1 <- cbind(mypd, mympd)

?mntd
#Explore the function and run the examples

mymntd <- mntd(t(mydata[,-c(1)]), cophenetic(mytree), abundance.weighted=T)
mymntd
myresults2 <- cbind(myresults1, mymntd)
colnames(myresults2) <- c("PD", "SR", "MPD", "MNTD")
myresults2
write.table(myresults2, file = "diversity_index_results.txt")

mycomdist <- comdist(t(mydata[,-c(1)]), cophenetic(mytree), abundance.weighted=T)
library(cluster)
mycomdist.clusters <- hclust(mycomdist)
plot(mycomdist.clusters, main = "Dendrogram based on beta MPD")

mycomdistnt1 <- comdistnt(t(mydata[,-c(1)]), cophenetic(mytree), abundance.weighted=F, exclude.conspecifics = T)
mycomdistnt1.clusters <- hclust(mycomdistnt1)
plot(mycomdistnt1.clusters, main = "Dendrogram based on beta MNTD excluding conspecifics")

mycomdistnt2 <- comdistnt(t(mydata[,-c(1)]), cophenetic(mytree), abundance.weighted=F, exclude.conspecifics = F)
mycomdistnt2.clusters <- hclust(mycomdistnt2)
plot(mycomdistnt2.clusters, main = "Dendrogram based on beta MNTD including conspecifics")


```


5. Finally we are going to check for phylogenetic clustering and overdispersion
```{r}
my.ses.pd <- ses.pd(t(mydata[,-c(1)]), mytree, null.model = "taxa.labels",
runs = 999)
my.ses.pd
write.table(my.ses.pd, file="ses.pd.result.txt")

my.ses.mpd <- ses.mpd(t(mydata[,-c(1)]), cophenetic(mytree), null.model = "taxa.labels",
abundance.weighted = T, runs = 999)
my.ses.mpd
write.table(my.ses.mpd, file="ses.mpd.result.txt")


my.ses.mntd <- ses.mntd(t(mydata[,-c(1)]), cophenetic(mytree), null.model = "taxa.labels",
abundance.weighted = T, runs = 999)
my.ses.mntd
write.table(my.ses.mntd, file="ses.mntd.result.txt")



comm.phylo.cor.r <- comm.phylo.cor(t(mydata[,-c(1)]), mytree, metric="cij",null.model="sample.taxa.labels")
comm.phylo.cor.r
par(mai=c(1.02,0.82,0.82,0.42))
hist(comm.phylo.cor.r$random.corrs)
abline(v=comm.phylo.cor.r$obs.corr)

```

6. Using means of each kind of community

```{r}

names(mydata)

Specie <- mydata$Specie 
Caliza <- rowMeans(mydata[c(2,3)])
Dolomia <- rowMeans(mydata[c(4,5)])
Brezal <- rowMeans(mydata[c(6,7,8,9)])
Alcornocal <- rowMeans(mydata[c(10,11,12,13)])
Quejigal <- rowMeans(mydata[c(14,15,16)])
mydata2 <- data.frame(Specie, Caliza, Dolomia, Brezal, Alcornocal, Quejigal)
mydata2



```



4. Calculate phylogenetic diversity statistics.

```{r}
color.plot.phylo(mytree, mydata2, "Caliza", "Specie", leg.cex = 0.6, main = "Caliza") 
color.plot.phylo(mytree, mydata2, "Dolomia", "Specie", leg.cex = 0.6, main = "Dolomia") 
color.plot.phylo(mytree, mydata2, "Brezal", "Specie", leg.cex = 0.6, main = "Brezal") 
color.plot.phylo(mytree, mydata2, "Alcornocal", "Specie", leg.cex = 0.6, main = "Alcornocal") 
color.plot.phylo(mytree, mydata2, "Quejigal", "Specie", leg.cex = 0.6, main = "Quejigal") 


?pd
#Explore the function and run the examples

rownames(mydata2) = mydata2[,1]
as.matrix(mydata2[,-c(1)])

mypd <- pd(t(as.matrix(mydata2[,-c(1)])), mytree, include.root=TRUE)
mypd

?mpd
#Explore the function and run the examples

mympd <- mpd(t(mydata2[,-c(1)]), cophenetic(mytree), abundance.weighted=T)
mympd
myresults1 <- cbind(mypd, mympd)

?mntd
#Explore the function and run the examples

mymntd <- mntd(t(mydata2[,-c(1)]), cophenetic(mytree), abundance.weighted=T)
mymntd
myresults2 <- cbind(myresults1, mymntd)
colnames(myresults2) <- c("PD", "SR", "MPD", "MNTD")
myresults2
write.table(myresults2, file = "diversity_index_results.txt")

mycomdist <- comdist(t(mydata2[,-c(1)]), cophenetic(mytree), abundance.weighted=T)
library(cluster)
mycomdist.clusters <- hclust(mycomdist)
plot(mycomdist.clusters, main = "Dendrogram based on beta MPD")

mycomdistnt1 <- comdistnt(t(mydata2[,-c(1)]), cophenetic(mytree), abundance.weighted=F, exclude.conspecifics = T)
mycomdistnt1.clusters <- hclust(mycomdistnt1)
plot(mycomdistnt1.clusters, main = "Dendrogram based on beta MNTD excluding conspecifics")

mycomdistnt2 <- comdistnt(t(mydata2[,-c(1)]), cophenetic(mytree), abundance.weighted=F, exclude.conspecifics = F)
mycomdistnt2.clusters <- hclust(mycomdistnt2)
plot(mycomdistnt2.clusters, main = "Dendrogram based on beta MNTD including conspecifics")


```


5. Finally we are going to check for phylogenetic clustering and overdispersion
```{r}
my.ses.pd <- ses.pd(t(mydata2[,-c(1)]), mytree, null.model = "taxa.labels",
runs = 999)
my.ses.pd
write.table(my.ses.pd, file="ses.pd.result.txt")

my.ses.mpd <- ses.mpd(t(mydata2[,-c(1)]), cophenetic(mytree), null.model = "taxa.labels",
abundance.weighted = T, runs = 999)
my.ses.mpd
write.table(my.ses.mpd, file="ses.mpd.result.txt")


my.ses.mntd <- ses.mntd(t(mydata2[,-c(1)]), cophenetic(mytree), null.model = "taxa.labels",
abundance.weighted = T, runs = 999)

my.ses.mntd
write.table(my.ses.mntd, file="ses.mntd.result.txt")



comm.phylo.cor.r <- comm.phylo.cor(t(mydata2[,-c(1)]), mytree, metric="cij",null.model="sample.taxa.labels")
comm.phylo.cor.r
par(mai=c(1.02,0.82,0.82,0.42))
hist(comm.phylo.cor.r$random.corrs)
abline(v=comm.phylo.cor.r$obs.corr)

```
